Maximizing Ink in Symmetric Partial Edge Drawings of $k$-plane Graphs

### 2018-01-08 eurocg/paper.tex
sec 4.2 alpha

\begin{theorem}
	Let $ G $ be a simple graph with $m$ edges and $ \Gamma $ a straight-line drawing of $ G $. If the intersection graph $ C = (V,E) $ of $ G $ is a tree with maximum degree $ k \in \mathbb{N} $ the \maxsped problem can be solved in $ O(mk + m \log m)) $ time and space.
\end{theorem}
\begin{proof}
	Since $ C $ is known to be a tree, and hence $ |E| = O(m) $, we can compute the intersection graph and the stub lengths in $ O(m\log m) $ time and space.

@@

\subsection{Cacti}
\label{sec:cactus}
In this section we restrict $ C = (V,E) $ to be a cactus graph, i.e. a simple graph in which no edge $ e \in E $ is part of more than one cycle. For a cactus graph $ C $ the \emph{block-cut tree} is the tree $ \mathcal T = (\mathcal V, \mathcal E) $where $ \mathcal V $ contains a vertex for every cycle and \emph{cut vertex} in $ C $. We call a vertex $ u \in \mathcal V $ a \emph{block vertex} if it represents a cycle and \emph{articulation vertex} else. For a block vertex $ u \in \mathcal V $ representing an induced cycle with vertices $ V_c \subseteq V $ we define $ \textit{cycle}(u) = V_c$. For an articulation vertex $ u \in \mathcal V $ representing the cut vertex $ v \in V $ we define $ A(u) = v $. There is an edge $ uv \in \mathcal E $ between a block vertex $ u $ and articulation vertex $ v $ whenever $ v \in \textit{cycle}(u) $.

@@

For an articulation vertex $ u \in \mathcal V $:
\begin{align}
\label{rec:cactus_block}
T_i(u) = \ell_i(A(u)) + \sum_{v \in c(u)}
\begin{cases}
\solshort(v) & \text{if } \ell_i(A(u)) \leq a \\
\solmid(v) & \text{if } a < \ell_i(A(u)) \leq b \\
\sollong(v) & \text{else}
\end{cases}
\end{align}

@@

\begin{lemma}
	\label{lem:compute_cactus}
	Let $ \mathcal T = (\mathcal V,\mathcal E) $ be a cactus with maximum degree $ k $ and vertex weights $ \ell_0,\dots,\ell_{\delta} $ for every articulation vertex $ u \in \mathcal V $ with degree $ \delta $. We can compute $ \solshort(u) $, $ \solmid(u) $ and $ \sollong(u) $ in $ O(\delta) $ time and space, if the values $ \sollong(v) $, $ \solmid(v) $ and $ \solshort(v) $ are computed for all children $ v \in c(u) $.
\end{lemma}
\begin{proof}
	\fabian{todo}
\end{proof}

\begin{theorem}
	Let $ G $ be a simple graph with $m$ edges and $ \Gamma $ a straight-line drawing of $ G $. If the intersection graph $ C = (V,E) $ of $ G $ is a cactus with maximum degree $ k \in \mathbb{N} $ the \maxsped problem can be solved in $ O(mk^2 + m \log m)) $ time and space.
\end{theorem}
\begin{proof}
	\fabian{todo}
\end{proof}

>>>>

\begin{theorem}
	\label{thm:tree}
	Let $ G $ be a simple graph with $m$ edges and $ \Gamma $ a straight-line drawing of $ G $. If the intersection graph $ C = (V,E) $ of $ G $ is a tree with maximum degree $ k \in \mathbb{N} $ the \maxsped problem can be solved in $ O(mk + m \log m)) $ time and space.
\end{theorem}
\begin{proof}
	Since $ C $ is known to be a tree, and hence $ |E| = O(m) $, we can compute the intersection graph and the stub lengths in $ O(m\log m) $ time and space.

@@

\subsection{Cacti}
\label{sec:cactus}
In this section we restrict $ C = (V,E) $ to be a cactus graph, i.e. a simple graph in which no edge $ e \in E $ is part of more than one cycle. For a cactus graph $ C $ the \emph{block-cut tree} is the tree $ \mathcal T = (\mathcal V, \mathcal E) $where $ \mathcal V $ contains a vertex for every cycle and \emph{cut vertex} in $ C $. We call a vertex $ u \in \mathcal V $ a \emph{block vertex} if it represents a cycle and \emph{articulation vertex} else. For a block vertex $ u \in \mathcal V $ representing an induced cycle with vertices $ V_c \subseteq V $ we define $ \textit{cycle}(u) = V_c$. For an articulation vertex $ u \in \mathcal V $ representing the cut vertex $ v \in V $ we define $ A(u) = v $. There is an edge $ uv \in \mathcal E $ between a block vertex $ u $ and articulation vertex $ v $ whenever $ v \in \textit{cycle}(u) $ and an edge $ uv \in \mathcal E $ between two articulation vertices whenever $ A(u)A(v) $ was an edge in $ E $.

@@

For an articulation vertex $ u \in \mathcal V $:
\begin{align}
\label{rec:cactus_articulation}
T_i(u) = \ell_i(A(u)) + \sum_{v \in c(u)}
\begin{cases}
\solshort(v) & \text{if } \ell_i(A(u)) \leq a \\
\solmid(v) & \text{if } a < \ell_i(A(u)) \leq b \\
\sollong(v) & \text{else}
\end{cases}
\end{align}

@@

\begin{lemma}
	\label{lem:compute_articulation}
	Let $ \mathcal T = (\mathcal V,\mathcal E) $ be a cactus with maximum degree $ k $ and vertex weights $ \ell_0,\dots,\ell_{\delta} $ for every articulation vertex $ u \in \mathcal V $ with degree $ \delta $. We can compute $ \solshort(u) $, $ \solmid(u) $ and $ \sollong(u) $ in $ O(\delta) $ time and space, if the values $ \sollong(v) $, $ \solmid(v) $ and $ \solshort(v) $ are computed for all children $ v \in c(u) $.
\end{lemma}
\begin{proof}
	Follows from Lemma~\ref{lem:compute_vertex}, since an articulation vertex has at most $ \delta $ many children in $ \mathcal{T} $, one for each cycle it is part of in $ C $ and one for every other articulation vertex it was already connected to in $ C $. 
\end{proof}

\begin{lemma}
	\label{lem:compute_block}
	Let $ \mathcal T = (\mathcal V,\mathcal E) $ be a cactus with maximum degree $ k $ and vertex weights $ \ell_0,\dots,\ell_{\delta} $ for every block vertex $ u \in \mathcal V $ with degree $ \delta $. We can compute $ \solshort(u) $, $ \solmid(u) $ and $ \sollong(u) $ in $ O(n\delta) $ time and space, if the values $ \sollong(v) $, $ \solmid(v) $ and $ \solshort(v) $ are computed for all children $ v \in c(u) $.
\end{lemma}
\begin{proof}
	Let $ G $, $ \Gamma $,  $ C = (V,E) $ and $ \mathcal T = (\mathcal V,\mathcal E) $ as above and $ u \in \mathcal V $ a block vertex with degree $ \delta $. Computing the value for the cycle itself $ \textit{cycle}(u) $ takes $ O(n) $ time by Bruckdorfer et al. Now at every vertex $ v \in \textit{cycle}(u) $ there might be a set of up to $ \delta $ many children, which are all articulation points. Their values are already known and maximization over them can hence be done in $ O(\delta) $ time by the same argument as in Lemma~\ref{lem:compute_vertex}. Following we need $ O(n\delta) $ to compute the three values $ \solshort(u) $, $ \solmid(u) $ and $ \sollong(u) $.
\end{proof}

\begin{theorem}
	Let $ G $ be a simple graph with $m$ edges and $ \Gamma $ a straight-line drawing of $ G $. If the intersection graph $ C = (V,E) $ of $ G $ is a cactus with maximum degree $ k \in \mathbb{N} $ the \maxsped problem can be solved in $ O(mk^2 + m \log m)) $ time and space.
\end{theorem}
\begin{proof}
	Follows as in Theorem~\ref{thm:tree} by maximizing over the tree using Lemma~\ref{lem:cactus_correctness} and Recurrence~\ref{rec:cactus_articulation}. Then the running time follows from constructing the intersection graph and the block-cut tree as well as Lemma~\ref{lem:compute_articulation} and~\ref{lem:compute_block}.
\end{proof}
